---
title: "General_Fungi_Sakhon_Alpha_Div"
output: html_document
date: "12 April 2025"
---

```{r warning=FALSE, message=FALSE}
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
```

###STEP1: Import ASV table for analysis using Phyloseq
```{r}
otus <- read.table("SakhonNakhonApril2025.final.txt",header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

###STEP2: Import taxonomy table for analysis using Phyloseq

```{r}
taxmat <- read.table("SakhonNakhonApril2025.taxonomy.fix.txt", header=T,sep="\t",row.names=1)
taxmat <- as(as.matrix(taxmat),"matrix")
TAX = tax_table(taxmat)
```


###STEP3: Import mapping file for analysis using Phyloseq

1.Check mapping file before import to R, R will automatically change sample name that starts with number or contain “-” in sample name. If you get error in this step, you should check sample name first. 

2.First column of first row should not start with #, R will not read the first row that starts with #

3. You can choose which samples to include in analysis by indicating specific group in the column

```{r}
meta = read.table("SakhonNakhonApril2025.mapping_file.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
```

Check if your metadata file has been import successfully and correctly, the output will show a table of your metadata file (mapping file). *If you do not have header, you might start your first row with # (remove # and reload your mapping file).

```{r}
head(meta)
```

Construct sample_data-class using imported metadata
```{r warning=FALSE}
sampleData <- sample_data(meta)
```


###STEP4: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData)
```

```{r warning=FALSE}
physeq
```

```{r}
physeq = subset_taxa(physeq, Kingdom == "Fungi")
physeq
```


```{r}
sample_data(physeq)$crop = factor(sample_data(physeq)$crop, levels = c('Rubber_tree_A', 'Rubber_tree_B', 'Cassava', 'Sugarcane'))

```


```{r}
alpha.diversity = estimate_richness(physeq, measures = c("Observed"))
data.anova = cbind(sample_data(physeq), alpha.diversity)
physeq.anova = aov(Observed ~ crop, data.anova)
summary(physeq.anova)
```

#Alpha diversity by condition

```{r}
physeq.plot.richness.condition = plot_richness(physeq, x = "crop", color=("crop"), measures=c("Observed")) + 
  geom_boxplot(lwd=0.5) + ggtitle("SakonNakhon Fungal Alpha Diversity by crop") + 
  stat_compare_means(method = "anova", label.y = 650, label.x = 0.65) + 
  theme_pubr(border= TRUE, legend = c("none")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_hue(l = 58) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) #+
  #geom_text(data=LABELS, aes(x=LABELS$zone, y = LABELS$ylabel , label = LABELS$Letters))

physeq.plot.richness.condition
```



#Statistic Alpha Diversity

```{r}
tukey.ps = TukeyHSD(x=physeq.anova, 'crop', conf.level = 0.95)

generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }

LABELS=generate_label_df(tukey.ps, 'crop')
names(LABELS) = c('Letters','crop')
```



```{r}
LABELS
```



```{r}
  ylabel <- data.frame("ylabel" = c(400, 620, 470, 620)) #LAC surface
#ylabel <- data.frame("ylabel" = c(340, 400, 340, 380, 290,  460, 510, 590, 460,  350)) #CLC surface
```



```{r}
LABELS$ylabel<-ylabel$ylabel
```



```{r}
physeq.plot.richness.crop = plot_richness(physeq, x="crop", color=("crop"), measures=c("Observed")) + 
  geom_boxplot(lwd=0.5) + ggtitle("SakonNakhon Fungal Alpha Diversity by crop") + 
  annotate("text", x = 0.98, y = 650, label = "paste('ANOVA, ', italic('p'), ' = 0.00077')", parse = TRUE, size = 3.5) +
  theme_pubr(border= TRUE, legend = c("none")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_hue(l = 58) +
  theme(axis.text.x = element_text(size=12, angle = , hjust = ), axis.text.y = element_text(size=12, angle = , hjust = ), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14)) + geom_text(data=LABELS, aes(y = LABELS$ylabel , label = LABELS$Letters))
physeq.plot.richness.crop
```


#Save to Tiff file
```{r warning=FALSE}
tiff("General fungal-plant Alpha diversity CROP.tif",units = "in",  width = 6.5, height = 4.5, res = 1000)
physeq.plot.richness.crop
dev.off()
```

#Save to Png file
```{r warning=FALSE}
png("General fungal-plant Alpha diversity CROP.png", units = "in",  width = 6.5, height = 4.5, res = 1000)
physeq.plot.richness.crop
dev.off()
```
