---
title: "Beta diversity PCoA data 2025"
author: "Niar Ibrahim"
date: "Updated 17 February 2025"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

###
This is documentation and codes for Fungal diversity analysis in NICE Project dataset.

###STEP1: Load all necessary packages for analysis
More information about Phyloseq can be found at the following link: [Phyloseq](https://joey711.github.io/phyloseq/)
If you get error in this step, you probably need to install any packages which causes error.


#Download and Install Packages
```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
library(ggforce)
library(microViz)
```


###STEP1: Import ASV table for analysis using Phyloseq
```{r}
otus <- read.table("D:/OneDrive/Dokumen/R/R data/SakonNakhon/SakhonNakhonApril2025.final.txt",header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

###STEP2: Import taxonomy table for analysis using Phyloseq
```{r}
taxmat <- read.table("D:/OneDrive/Dokumen/R/R data/SakonNakhon/SakhonNakhonApril2025.taxonomy.fix.txt", header=T,sep="\t",row.names=1)
taxmat <- as(as.matrix(taxmat),"matrix")
TAX = tax_table(taxmat)
```

###STEP3: Import mapping file for analysis using Phyloseq

1.Check mapping file before import to R, R will automatically change sample name that starts with number or contain “-” in sample name. If you get error in this step, you should check sample name first. 

2.First column of first row should not start with #, R will not read the first row that starts with #

3. You can choose which samples to include in analysis by indicating specific group in the column

```{r}
meta = read.table("SakhonNakhonApril2025.mapping_file.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
```


Construct sample_data-class using imported metadata
```{r warning=FALSE}
sampleData <- sample_data(meta)
```


```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData)
physeq
```


```{r}
physeq <- subset_taxa(physeq, Kingdom == "Fungi")
physeq
```


#Selecting the WHO FPPL
```{r }
#critical group #non of the critical fungi was detected
#physeq.1 = subset_taxa(physeq, Species == "Cryptococcus neoformans")
#physeq.2 = subset_taxa(physeq, Species == "Candida auris")
physeq.3 = subset_taxa(physeq, Species == "Aspergillus fumigatus")
#physeq.4 = subset_taxa(physeq, Species == "Candida albicans")
#physeq.100 = subset_taxa(physeq, Genus == "Candida")

#High group #only candida tropicalis
#physeq.5 = subset_taxa(physeq, Species == "Nakaseomyces glabrata")
#physeq.6 = subset_taxa(physeq, Genus == "Histoplasma")
#physeq.7 = subset_taxa(physeq, Species == "Candida tropicalis")
#physeq.8 = subset_taxa(physeq, Species == "Candida parapsilosis")
#physeq.17 = subset_taxa(physeq, Genus == "Madurella")
#physeq.18 = subset_taxa(physeq, Species == "Falciformispora senegalensis")
#physeq.19 = subset_taxa(physeq, Species == "Curvularia lunata")
#physeq.20 = subset_taxa(physeq, Species == "Zopfia rosatii")
physeq.21 = subset_taxa(physeq, Genus == "Acremonium")
physeq.22 = subset_taxa(physeq, Genus == "Fusarium")

#Mucorales
#physeq.23 = subset_taxa(physeq, Order == "Mucorales")
physeq.23 = subset_taxa(physeq, Genus == "Rhizopus")
#physeq.24 = subset_taxa(physeq, Genus == "Mucor")
#physeq.25 = subset_taxa(physeq, Genus == "Lichtheimia")

#Medium group
physeq.9 = subset_taxa(physeq, Genus == "Scedosporium")
#physeq.10 = subset_taxa(physeq, Species == "Lomentospora prolificans")
#physeq.11 = subset_taxa(physeq, Genus == "Coccidioides")
#physeq.12 = subset_taxa(physeq, Species == "Pichia kudriavzevii")
#physeq.13 = subset_taxa(physeq, Species == "Cryptococcus gattii")
#physeq.14 = subset_taxa(physeq, Species == "Talaromyces marneffei")
#physeq.15 = subset_taxa(physeq, Species == "Pneumocystis jirovecii")
#physeq.16 = subset_taxa(physeq, Genus == "Paracoccidioides")
```


```{r}
# Critical group: Candida albicans - High group: Acremonium, Fusarium, Mucorales - Medium group: Scedosporium, Pichia kudriavzevii
physeq.merge = merge_phyloseq(physeq.3, physeq.9, physeq.21, physeq.22, physeq.23)

physeq.merge
```


```{r}
sample_data(physeq.merge)$crop = factor(sample_data(physeq.merge)$crop, levels = c('Rubber_tree_A', 'Rubber_tree_B', 'Cassava', 'Sugarcane'))

```


```{r}
# 🔹 Keep only taxa that appear in at least one sample (non-zero total abundance)
physeq.merge <- prune_taxa(taxa_sums(physeq.merge) > 0, physeq.merge)
```



```{r}
# ----------------------------
# Alternative: Aitchison CLR + PCoA
# ----------------------------
# 1️⃣ CLR Transformation (Aitchison)
# ----------------------------
physeq.clr <- phyloseq::transform_sample_counts(physeq.merge, function(x) x + 1) %>%
  microViz::tax_transform("clr", rank = "Species")
```


```{r}
# ----------------------------
# 2️⃣ PCoA Ordination (Euclidean)
# ----------------------------
ord_clr <- ordinate(physeq.clr, method = "PCoA", distance = "euclidean")
```


```{r}
# ----------------------------
# 3️⃣ Prepare data for plotting
# ----------------------------
plot_df <- as.data.frame(ord_clr$vectors)
plot_df$SampleID <- rownames(plot_df)
```


```{r}
meta_df <- data.frame(sample_data(physeq.clr))
meta_df$SampleID <- rownames(meta_df)
```


```{r}
# Clean duplicates and merge
meta_df_clean <- meta_df[, !duplicated(colnames(meta_df))]
plot_df <- merge(plot_df, meta_df_clean, by = "SampleID", all.x = TRUE)
```

```{r}
plot_df$crop <- plot_df$crop
```

```{r}
colnames(plot_df)
```

```{r}
# ✅ Fix crop column
# (choose the correct column that exists — likely crop.x or crop.y)
if ("crop.x" %in% colnames(plot_df)) {
  plot_df$crop <- plot_df$crop.x
} else if ("crop.y" %in% colnames(plot_df)) {
  plot_df$crop <- plot_df$crop.y
} else {
  stop("No crop column found! Check your metadata column names.")
}
```


```{r}
# ----------------------------
# 4️⃣ PERMANOVA (Aitchison distance)
# ----------------------------
dist_euc <- phyloseq::distance(physeq.clr, method = "euclidean")
set.seed(123)
permanova <- adonis2(dist_euc ~ crop, data = plot_df)
```


```{r}
p_value <- format(permanova$`Pr(>F)`[1], digits = 3)
```


```{r}
# ----------------------------
# 5️⃣ PCoA Plot with geom_mark_hull and p-value
# ----------------------------
p_clr <- ggplot(plot_df, aes(Axis.1, Axis.2, color = crop)) +
  geom_point(size = 2.5, alpha = 0.9) +
  geom_mark_ellipse(
  aes(fill = crop, group = crop),
  alpha = 0.2,
  color = NA,
  label.fontsize = 9,
  label.fontface = "bold",
  label.buffer = unit(0.25, "lines"),
  label.fill = "white",
  label.colour = "black",
  show.legend = FALSE
) +
  ggtitle("PCoA (Aitchison CLR) of WHO FPPL 11 species" ) +
  annotate(
    "text",
    label = "bold(paste('PERMANOVA, ', italic('p'), ' = 0.002'))", 
    parse = TRUE, 
    size = 3,
    hjust = 0,
    vjust = 1,
    x = min(plot_df$Axis.1) * 1.5,
    y = max(plot_df$Axis.2) * 2.2,
  ) +
  coord_equal() +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "right",  # move legend inside bottom-right
    legend.background = element_rect(fill = "white", color = NA),
    legend.key.size = unit(0.6, "cm"),
    legend.title = element_text(face = "bold", size = 11),
    legend.text = element_text(size = 10),
    plot.title = element_text(size = 12, hjust = 0.5, vjust = 1),
    axis.title = element_text(face = "bold", size = 13),
    axis.text = element_text(size = 11, color = "black"),
    plot.margin = margin(5, 5, 5, 5),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.4)
  )

p_clr
```

#Save to jpeg file
```{r warning=FALSE}
jpeg("D:/OneDrive/Dokumen/R/R data/SakonNakhon/Abundance - Distribution/WHOFPPL Beta diversity PCoA-Species-Aitchison SakonNakhon 2025.jpeg",units = "in",  width = 6, height = 6, res = 1000)
p_clr
dev.off()
```


```{r}
# Convert to dataframe
clr_df <- speedyseq::psmelt(physeq.clr)

# Count unique species per crop
clr_species_per_crop <- clr_df %>%
  dplyr::group_by(crop) %>%
  dplyr::summarize(n_species = dplyr::n_distinct(Species))

clr_species_per_crop
```


```{r}
ntaxa(physeq.clr)
```